### 2016-05-03 paper/paper.tex
add background, cleanup things

\input{headers}

\begin{abstract}

Since its creation in 2009, Bitcoin has used a hash-based proof-of-work to
generate new blocks, and create a single public ledger of transactions. The
hash-based computational puzzle employed by Bitcoin is instrumental to its
security, preventing sybill attacks and making double-spending attacks more
difficult. However, there have been concerns over the efficiency of this
proof-of-work puzzle, and alternative ``useful'' proofs have been proposed.

In this paper, we present DDoSCoin, which is a cryptocurrency with a
\emph{mallicious} proof-of-work. DDoSCoin allows miners to prove that they have
contributed to a distributed denial of service attack against specific target
servers.  This proof involves making a large number of TLS connections to a
target server, and using infrequent cryptographic responses to prove that a
large number of connections has been made. Like proof-of-work puzzles, these
proofs are inexpensive to verify, and can be made arbitrarily difficult to
solve.

\end{abstract}

\section{Introduction}

Cryptocurrencies rely on proofs-of-work (PoW) that
require miners to spend a large amount of effort to solve a specific puzzle,
such that once solved, allowing others to inexpensively verify the solution is a
valid one. In Bitcoin, the proof-of-work is a computational puzzle based on the
SHA256 hash function, and miners are tasked with finding a partial preimage. The
only known way to do this is to iterate over a large number of inputs to the
hash function, and check if the output hash is less than a specified target. In
simplified terms, to find an input that hashes to N bits of 0s, the miner must
perform on average $2^{N}$ hashes. However, once found, other miners and Bitcoin
nodes can verify the puzzle solution with
a single hash~\footnote{Bitcoin uses double-SHA256 in its design, but this
detail is not important for the purposes of this paper.}.

Although the proof-of-work used in Bitcoin gives it resistance to sybil attacks,
the amount of computational effort carried out collectively by its miners does
not contribute to any problems besides securing the currency from attack.
Indeed, this computational effort is substantial: As of 2016, miners
collectively perform roughly $10^{18}$ (or about $2^{60}$) hashes every
second~\cite{blockchain}.

Previously, researchers have proposed alternative cryptocurrencies (altcoins)
that aim to have more beneficial proofs-of-work~\cite{primecoin,permacoin} that
provide utility beyond securing the underlying currency. In this paper however,
we investigate going in the opposite direction, and propose an altcoin that has
a \emph{malicious} proof-of-work that is externally detrimental.

In particular, we propose a proof-of-work that allows miners to prove they have
participated in a distributed denial of service (DDoS) attack against a
particular target. Miners are incentivized to send and receive large amounts of network
traffic at the target in order to produce a valid proof-of-work. As in other
cryptocurrencies, these proofs can be inexpensively verified by others, and the
original miner can collect a reward. This reward can be sold for other
currencies (including Bitcoin or even traditional currencies), allowing botnet
owners and other attacks to directly collect revenue for their assistance
in a decentralized DDoS attack.

The malicious proof-of-work operates by having miners create a large number of
TLS connections to a target webserver, and using the server's signed responses
as a proof of connection. In modern versions of TLS, the server signs a
client-provided parameter of the connection, along with server-provided values
used in the key exchange of the connection. This allows the client to prove to
others that it has communicated with the server. In addition, the signed value
returned by the server is not predictable to the client, and is randomly
distributed. Thus, clients can use a similar trick as in Bitcoin, and only
report connections that match some rare threshold, such as the signature from
the server starts with $N$ bits of 0s. On average, it will take clients $2^{N}$
connections to produce such a proof.

Although the malicious proof-of-work only works against websites that support
TLS 1.2, as of April 2016, over 56\% of the Alexa top million websites support
this version of TLS~\cite{censys}. Furthermore, we expect this number to
increase as TLS support becomes more widespread~\cite{letsencrypt}.

\emph{Contributions}
\begin{itemize}
\item Propose a novel conceptual cryptocurrency, DDoSCoin, whose proof-of-work
incentives miners to participate in a DDoS attack, and prove they have done so.
\item Implement our proof-of-work function and evaluate its performance.
\item Discuss and estimate the impact such a cryptocurrency would have if
deployed, and outline defenses.
\end{itemize}

Section blah does blah, and Section blat describes blat.

\section{Related Work}

Ethereum
Permacoin
Primecoin
Other interesting proofs-of-work?

Botnet-as-a-service

\section{Background}

In this section, we review relevant basics of cryptocurrency. For a more
detail, we refer the reader to other
sources~\cite{bitcoin-book,bitcoin-paper,bitcoin-sok}.

\subsection{Cryptocurrency}

Bitcoin employs a hash-based proof-of-work protocol to construct a public ledger
of transactions. In this protocol, miners attempt to solve a simple hash-based
computational puzzle in order to mine blocks. Mining a block nets the miner a
reward, and the right to specify what new (valid) transactions should be
included in the public ledger. To mine a block, a miner must find a hash of a
block header that is less than an adjustable target. The block header includes
the hash of the last block found, the merkle root of the list of transactions
the miner wants to include in this block, and a nonce. The easiest way to solve
this puzzle is to iterate the nonce\footnote{or modify transactions, therby
changing the merkle root} until the hash satisfies the target difficulty. Then,
the miner publishes the found block, and miners use this block as the latest
block.

\subsection{Proof-of-Stake}

Proof-of-Stake is an alternate way of generating new blocks in a cryptocurrency.
In this system, blocks are ``minted'' (rather than mined) based on how much
stake the miner has in the currency (rather than their fraction of computational
power). Proof-of-stake assumes that the underlying cryptocurrency is already
somewhat distributed, for example through the use of a traditional proof-of-work
system. In this subsection, we will describe Peercoin, an active
alternate cryptocurrency that employs a hybrid proof-of-work and proof-of-stake
algorithm to mint blocks. For more details on proof-of-stake, we refer the
reader to the Peercoin~\cite{peercoin} and NXT~\cite{nxt} whitepapers.

Peercoin's proof-of-stake allows currency holders to mint new blocks based on
how many coin-days they are able and willing to consume. ``Coin-days
consumed'' is a measure of the product of the amount of coins spent, and the
number of days the have not been previously spent. For example, to consume 10
coin-days, a currency holder can spend 10 Peercoin that they have held for a
day, or equivalently, spend 5 Peercoins that were last spent 2 days ago. Note
that the holder can send these coins to anyone (including themselves) in order
to consume the coin age.

To mint a new block, a minter creates a special transaction (presumably one
where the minter pays themself) where one of the
inputs has been unspent for at least 30 days. The resulting transaction, along
with a timestamp, must meet a certain hash target that gets easier proportional to how many
coin-days are consumed. For example, if the current difficulty gives a minter
consuming 100 coin-days a 1\% chance of minting a block per second, a minter
that was able to consume 200 coin-days would have a 2\% chance per second.

The input to the hash includes information about the ``aged'' transaction, as
well as a current timestamp\footnote{In recent versions of Peercoin, the hash
also includes a stakeModifier which is computed from recent blocks; however,
these details are not important for our purposes}. If this hash is less than the
current target multiplied by the number of coin-days consumed in the
transaction, then the transaction meets the proof-of-stake difficulty and a new
block is minted. Although this looks similar to a proof-of-work check where a
hash is being compared to a target, here the only varying inputs are the
long-lived (aged) input transaction, and a corse-grained timestamp with a
resolution of 1 second. Thus, for each aged unspent transaction output they
control, minters perform only one hash per second in order to solve the
proof-of-stake.

% hash(nStakeModifier + txPrev.block.nTime + txPrev.offset + txPrev.nTime +
%       txPrev.vout.n + nTime) < bnTarget * nCoinDayWeight
%
% ss << nTimeBlockFrom << nTxPrevOffset << txPrev.nTime << prevout.n <<
%   nTimeTx;
% given a new tx that is going to be my "proof-of-stake" tx,
% which has an input from prevTx (where prevTx is at least 30 days old)
% H(nStakeModifier + txPrev.block.nTime + txPrev.offset + txPrev.nTime +
%   txPrev.vout.n + tx.nTime)

% stakeModifiers (in v0.3): you start from the block where txPrev was in,
% and then walk forward until the first block that was made more than
% GetStakeModifierSelectionInterval() seconds after that the block txPrev was in (about 8.8 days...)
% If that block has the stakeModifier flag set, use its stakeModifier

% Ist das system relevant?
In Peercoin, coins cannot age beyond 90-days; coins that have not been consumed
after 90-days are counted as 90-day-old coins. This prevents very old
coin-holders from being able to disrupt the blockchain at a future date,
potentially causing a large fork of the blockchain and destabalizing it.

When a proof-of-stake is minted, the minter is allowed to collect a small
subsidy proportional to the coin-age consumed in their proof-of-stake
transaction. This serves as the only reward for a minter, as transaction fees are not
collected by minters in Peercoin.

TODO: how does Peercoin make sure other mint?ers can't steal the coinstake
transaction? Is it signed by the coinstake kernel, or does the coinstake kernel
include the merkle root of transactions somehow?

\section{DDoSCoin Design}

Miners in DDoSCoin repeatedly create connections to a TLS victim server, and
check for a response that satisifes a target difficulty decided by the network.
If the response satisifies this condition, then parameters of the TLS handshake
can be published by the miner to create a new valid block.

To begin, a miner generates a random 32-byte secret nonce, $N$. The miner also
selects the latest block in the chain it is mining on, and the set of
transactions that the miner intends to include in this block, both as is done in
Bitcoin. The miner then takes the SHA256 hash of the previous block header, the
Merkle root of transactions (as generated in Bitcoin), and the secret nonce $N$,
and hashes their concatenation using SHA256 to generate the
32-byte \texttt{client\_random} the miner will use.

The miner initiates a TLSv1.2 connection with the victim server, using the
\texttt{client\_random}, and choosing a set of cipher suites that will ensure
the server will send a server key exchange message. This includes ephemeral
Diffie-Hellman (DHE), and ephemeral elliptic curve Diffie-Hellman (ECDHE). We
note that for the default cipher suite list sent by Google Chrome, over 94\%
of TLS 1.2 servers in the Alexa top million choose a cipher suite with a server key exchange
message~\cite{censys}, compatible with DDoSCoin.
% https://censys.io/domain/report?q=443.https.tls.version%3A%22TLSv1.2%22&field=443.https.tls.cipher_suite.name&max_buckets=

The server will respond with a server hello message, containing a 32-byte
\texttt{server\_random}, the cipher suite that will be used, and other
parameters that are not needed for the purposes of DDoSCoin.

Next, the server sends its certificate chain, and a server key exchange message.
The certificate chain contains the server's (signed) TLS certificate, as well as
intermediate certificate authorities (CAs) that create a signature chain to a
browser-trusted root CA. Inside the server's certificate is a public
key, as well as the identity of the server (i.e. it's domain). The miner stores
these for verification in case a block is found.

In the server key exchange message, the server sends key exchange parameters
(such as the Diffie-Hellman parameters and the ephemeral public value generated
by the server, or the curve parameters and a public curve point in ECDHE). These
key exchange parameters, along with the \texttt{client\_random} and
\texttt{server\_random},  are signed by the server using its private key; this
signature can be verified by using the public key in the certificate.

Because the server key exchange signature contains the \texttt{client\_random}
and \texttt{server\_random}, the signature value will be unpredictable to the
client ahead of time. In other words, each signature will be a random value that
depends on miner-provided value. Thus, it can be used as a proof-of-work in the
following way:

If the SHA256 hash of the server key exchange parameters, signature, and
miner-chosen nonce $N$ is less than the current target difficulty, then this
connection satisfies the proof-of-work threshold, and can be used to form the
next block.

To make a block, the miner publishes the prevoius block's hash, the transactions
(and their Merkle root), the nonce $N$, the \texttt{server\_random} value, the
server's certificate chain, the server key exchange parameters, and the server
key exchange signature.

\subsection{Validating Blocks}

To verify that a block is valid, miners must validate several items. First, they
must validate that the previous block hash and Merkle roots are valid, as in
Bitcoin. Next, they must re-create the \texttt{client\_random}, by hashing the
previous block hash, transaction Merkle root, and provided nonce. Then, they
must validate the certificate actually belongs to a valid victim server. This is
done by checking that the certificate chain is valid, rooted in a trusted
certificate authority, and the domain name is in a list of acceptable victims.
This list of valid victim domains is generated by concensus, and details for its
generation are discussed in Section~\ref{sec:victim}.

After the victim has been verified to be a valid one, the public key is used to
verify the server key exchange signature. This signature is over the
\texttt{client\_random}, \texttt{server\_random}, and server key exchange
parameters.

Finally, the block must be verified to meet the current target difficulty. The
server key exchange parameters, signature, and the block's nonce ($N$) are
hashed using SHA256, and compared to the current target difficulty. If it is
less than the target, the block is a valid one and becomes the latest block in
this chain.

\subsection{Earlier TLS Versions}

DDoSCoin is incompatible with earlier TLS (or even SSL) versions because the
server key exchange signature does not include the \texttt{client\_random}, or
any client-provided values. If we were to try to base a proof-of-work off of an
earlier version of TLS, the only value that provably comes from the server (i.e.
is signed by a key tied to the victim's identity) does not contain any
comittments from the client about the previous block, or what transactions
should be included in the block. If we were to accept blocks where there was
only a server signature that met some target difficulty, any miner could steal
another miner's found block (and the contained block reward) by changing the
transactions and forwarding the block.

If we were to somehow tie the previous block's hash and Merkle root into this
proof, for example, by saying that the hash of the signature and a
client-provided nonce had to be less than some target value, the proof-of-work becomes
a computational puzzle: a miner only has to collect a single signature from
the server, and try different nonces until the hash of signature and nonce meets
the target difficulty.

Similarly, proof-of-work target difficulties can not be based off of the
purported shared secret that results from a connection, because the client can
always choose different key exchange values to ensure that the resulting shared
secret is less than than the threshold difficulty, resulting in a computational
puzzle. Even values sent by the server---such as the Finished message that
contains a MAC dependent on the shared secret---can be (re)computed by the client.

This interesting property gives rise to an observation that with respect to
clients, the contents of TLS connections are \emph{deniable}. That is, it is not
possible for a client to record a TLS connection, including its chosen secrets,
and report this to a third party in a way that cryptographically proves what the
server says. This property has been accomplished in different more efficient ways in various
secure messaging protocols such as Off-the-Record (OTR)~\cite{otr}.

Besides TLS 1.2, there are other cryptographic protocols that have the feature
of the server signing a client-chosen value. For example, SSH ... \TK

Although not yet implemented, the TLS 1.3 proposal appears to also have ... \TK

\subsection{Victim Selection}
\label{sec:victim}

If DDoSCoin accepted proofs against any TLS 1.2 server, miners would be
incentivized to set up their own TLS servers locally, and focus exclusively on
``attacking'' those servers instead of remote ones. Since the miner would have
access to the private key for that server, they would not be required to even
create network connections, leading to the proof-of-work being a computational
puzzle rather than one based on participation in a network denial of service.
To prevent this, DDoSCoin must have a way to agree on what victims are
acceptable targets for the proofs-of-work to be considered valid. Similarly, it
is important to focus the denial-of-service effort toward a single or small
number of victims in order to have the most impact.

One simple solution is to hardcode a set of victim certificates into the
protocol.  For example, DDoSCoin's published code could contain the identities
of Facebook's servers and domains, and only allow and incentivize attacks
against them. However, there are many attackers that might feel uncomfortable
attacking Facebook, but would be willing to participate in attacks against other
websites. In addition, having only a single or small set of victim websites is a
risk to DDoSCoin: if those websites all go offline or successfully mitigate the
attacks (such as by disabling TLS), no future DDoSCoin blocks can be mined, and
the currency will stall. At this point, all DDoSCoins in circulation would
become nearly worthless, as there would be no way to safely transact them.

These issues lead to a few subtle requirements for choosing valid victims.
First, there must be \emph{limits} on who is allowed to be a victim, otherwise
miners can create local computational puzzles. At the same time, there must be
\emph{enough} targets in order to be robust against stalled blockchains.
Similarly, the list of victims must be \emph{flexible} in order to allow
removing targets as they become impractical to attack (either
through successful attack or mitigation), and adding new ones to take their
place.

To address these issues, we propose two mechanisms the DDoSCoin protocol can
employ to allow victims to be selected by concensus.

\subsubsection{Pay-to-DDoS}

In order to allow victims to be (temporarily) selected for DoS, DDoSCoin allows
``bounties'' for targetting specific servers. To accomplish this, DDoSCoin
introduces a new payment opcode, PAY\_TO\_DDOS, that can be used in transactions
subject to certain constraints. This opcode takes two arguments in an output
script: a string representing a domain name that the payer wishes to have
DDoSed, and a target difficulty corresponding to the amount of connections the
payer wishes to be made. Once this transaction is stored in a valid block,
anyone can collect its reward by creating a connection to the specified victim
server and obtaining a response that meets the given target difficulty. These
connections are made similar to the way that miners connect to victim servers to
create blocks, but with a few important differences. First, the
\texttt{client\_random} is generated using the PAY\_TO\_DDOS transaction ID
(txid), the SHA256 hash of the output script the payee wants to collect the
payment with, and a random nonce.  The client then makes connections to the
specified victim domain until it gets a response that meets the target
difficulty specified by the PAY\_TO\_DDOS transaction.
(SHA256(\texttt{server\_key\_exchange\_params} || \texttt{signature} ||
\texttt{nonce}) < target).

To collect the bounty reward, the client creates an input script that will
satisfy the PAY\_TO\_DDOS output. This includes the parameters of the
connection: the nonce, the \texttt{server\_random} value, the server's
certificate chain, the server key exchange parameters, and the server key
exchange signature. To validate the transaction, a miner must recompute the
\texttt{client\_random} using the transaction ID of the PAY\_TO\_DDOS
transaction, the hash of the output script in the claiming transaction, and the
nonce. Then, the miner must verify that the certificate is valid for the
provided domain (and rooted in a trusted CA), and that the certificate's
signature over the server key exchange message is valid. Finally, the server
must check that the server's response meets the target difficulty. If it does,
then the transaction is valid and can be included in a block.

There are a couple important features of this transaction. First, the reason the
transaction ID of the PAY\_TO\_DDOS transaction is included (committed to) in
the \texttt{client\_random} is to prevent a client from ``double-claiming''
multiple bounty rewards with the same proof-of-DoS. Similarly, the client
commits to (a hash of) the output script of the claiming transaction to
prevent others from being able to swap it for their own, therby ``stealing'' the
proof-of-DoS and collect the reward themselves. These properties combined allow two
mutaully-distrusting parties to directly and safely transact currency in
exchange for a specified-amount of work in a DoS attack.

% Maybe move this to discussion....
Since PAY\_TO\_DDOS transactions can be claimed by anyone, there is a risk to
the collector that by the time they complete their DoS attack, someone else will
have collected the payment. In addition, a potential victim website could
disincentivze attacks by submitting PAY\_TO\_DDOS transactions that they can
quickly solve locally. This would cause attackers to try to connect to them, and
eventually even collect a response that meets the target difficulty. However,
when the attacker went to publish this to collect the reward, the victim could
publish their proof, essentially double-spending the PAY\_TO\_DDOS transaction.
With some probability, the victim's transaction will ``win'' and prevent the
attacker from collecting the reward. Over the long term, attackers may be wary
of attacking such targets, and instead require higher payouts for attacking them
to cover such risks.

% client_random = H(txid || output_script || nonce)

%The transaction has no other parameters, but its value must be
%greater than the coinbase reward a miner would normally receive, and it must
%carry a non-negligble transaction fee. Once this transaction is stored in a
%valid block, the specified domain name is temporarily added to the list of
%allowed victims.  The first miner to mine a block with that domain as the victim
%must specify the PAY\_TO\_DDOS transaction as an input to its coinbase
%transaction (with any arbitrary output address), and have no other coinbase
%transactions. In other words, the miner collects the value in the PAY\_TO\_DDOS
%transaction in lieu of the normal block reward.
%
%The observant reader will notice that nothing stops the payer of the
%PAY\_TO\_DDOS transaction and the miner that it goes to from being the same
%entity. Indeed, a miner could create a PAY\_TO\_DDOS transaction with a victim
%domain that only it has access to, ensuring that it will be able to mine the
%block without competition, and collect its own payment. However, this will not
%net the miner new coins (\TK beyond any transaction fees collected? But we could
%make this block contain no transactions), as the coinbase transaction is
%replaced with its own PAY\_TO\_DDOS input.

While this new payment type allows a one-time payment for attacking a particular
target, it cannot be used to update the long-term victim list used by miners to
decide valid victims for blocks. Otherwise, miners could make a PAY\_TO\_DDOS
transaction specifying their own local domain in order to always have a
local-computation puzzle that allowed them to create new blocks in the future.
Instead, a second method for choosing victims must be employed.

%Instead, PAY\_TO\_DDOS must be a one-time only payment, allowing for the
%creation of a single block with a payment greater than the coinbase reward. This
%is similar to proof-of-burn, where new blocks can be created
%by miners proving that they have destroyed a certain amount of currency. The
%amount of currency in proof-of-burn cryptocurrencies is adjusted in the same way
%that difficulty is adjusted for proof-of-work to maintain a constant block rate.
%
%\paragraph{Concerns} Because it is possible for a miner to effectively pay a
%small fee to create a block that only they can solve quickly, PAY\_TO\_DDOS
%could irronically create a denial of service potential for DDoSCoin. To mitigate
%the risk of miners using PAY\_TO\_DDOS as a way to keep other miners from
%producing blocks, the amount above the coinbase reward a PAY\_TO\_DDOS
%transaction must pay is adjusted dynamically so that blocks that collect these
%transactions are not able to happen frequently. Additionally, these blocks do
%not contribute to the

\subsubsection{Alternate Proof-of-Work}

In order to allow the victim list to be updated dynamically, DDoSCoin uses
\emph{Proof-of-Stake} to let currency holders make small changes to the list of
acceptable victim domains. Using proof-of-stake ensures that updates are only
done by actors that already have a substantial stake in the currency, either by
purchasing it or by previously participating in the (DoS) mining themselves.

To update the list, a currency holder ``mints'' a proof-of-stake transaction
similar to Peercoin. First, they must hold a number of coins that have been
unspent for more than 30 days. After the 30 days have passed, each second the
minter
hashes the aged transaction, a recent stake-modifier block, and a current
(seconds-resolution) timestamp. If this hash is less than the current
proof-of-stake target difficulty multiplied by the coin-age being consumed in
this transaction, then the currency holder is allowed to make a new
proof-of-stake block. Inside this block, the currency holder can make one
addition to (or one deletion from) the current victim list.

Proof-of-stake blocks contains no coinbase or coinstake reward, and transaction
fees are not collected. Thus, there is no direct short-term monetary incentive
to produce such a block. The intention is that this will encourage only those
with a vested interest in the long-term success of DDoSCoin to participate in
updating the victim list.

After this block has been published, miners must wait until 10 proof-of-DoS
(non-proof-of-stake) blocks have been mined before the specified delta to the
change list is applied. This discourages malicious miners from forking the chain
at an earlier block with a proof-of-stake block adding their local domain to the
victim list, and immediately mining blocks on top of it to catch up to the main
chain.

%Proof-of-Stake to decide a delta on the victim-list.

%-However, the victim could easily generate a new certificate; should instead
%base identity off domain rooted to valid root CAs.

%DDoSCoin must select a small set of victim servers to p

%-Miners vote (proof-of-stake-esq)

\section{Discussion}

-Proof of (burnt) bandwidth
-Defenses
-Target server's participation

\subsection{Ethical Considerations}
%¯\\\_(ツ)\_/¯
%\symbol{2544}
\shrug

% ¯\_(ツ)_/¯

\section{Conclusion}

{\footnotesize\balance
\bibliographystyle{plain}
\bibliography{paper}}

\end{document}