### 2016-05-08 paper/figures.tex
add btc-tx

\newcommand{\FigTx}{
\begin{figure*}[ht]
    \centering
    \includegraphics[width=0.9\textwidth,trim={0 6.5cm 1cm 0},clip]{btc-transaction.pdf}
    \caption{Standard Bitcoin transaction script. Transaction \#10 spends \#9 by providing a
            valid input script that when prepended to Transaction \#9's output script,
            executes to produce a valid output. The output script stores a hash of a
            public key (verified using the OP\_DUP OP\_HASH160 and OP\_EQUALVERIFY opcodes),
            and requires a valid signature over the spending transaction (verified via
            OP\_CHECKSIG).}
\end{figure*}
}